---
title: "Growth Modeling Notes"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
df <- read.csv('data.csv')
```

## Data Prelims For Growth Modeling Using SEM

* Requires wide rather than long data

* More notes about that here...

* Change `df` to wide using the "reshape" package.

```{r}
library(reshape2)
df_wide <- reshape(df, idvar = 'id', timevar = 'time', direction = 'wide')

```

## The Basic Growth Model

Some notes on this section.

* Point one
* Another point
* Another important note

An equation or two...

\begin{equation}
y_{it} = b_0 + b_1 t + e_t
\end{equation}

### Growth Model Syntax Notes

Latent intercept term
* constrain each factor loading to one

```{r, eval = F}
latent_int =~ 1*happy.1 + 1*happy.2 + 1*happy.3 + 1*happy.4 + 1*happy.5 + 1*happy.6
```

Latent slope term
* start from zero and increase each basis coefficient by one

```{r, eval = F}
latent_slope =~ 0*happy.1 + 1*happy.2 + 2*happy.3 + 3*happy.4 + 4*happy.5 + 5*happy.6
```

Estimate the mean and variance of both latent terms

```{r, eval = F}
latent_int ~~ latent_int
latent_slope ~~ latent_slope
```


...continue with notes

Here is the full model:

```{r}
library(lavaan)

linear_change_string <- '

# Latent intercept and slope factors

intercept_happy =~ 1*happy.1 + 1*happy.2 + 1*happy.3 + 1*happy.4 + 1*happy.5 + 1*happy.6
slope_happy =~ 0*happy.1 + 1*happy.2 + 2*happy.3 + 3*happy.4 + 4*happy.5 + 5*happy.6

# Mean and variance of latent factors

intercept_happy ~~ intercept_happy
slope_happy ~~ slope_happy

# Covariance between latent factors

intercept_happy ~~ slope_happy

# Fix observed variable means to 0

happy.1 ~ 0
happy.2 ~ 0
happy.3 ~ 0
happy.4 ~ 0
happy.5 ~ 0
happy.6 ~ 0

# Constrain residual (error) variance of observed variables to equality across time

happy.1 ~~ res_var*happy.1
happy.2 ~~ res_var*happy.2
happy.3 ~~ res_var*happy.3
happy.4 ~~ res_var*happy.4
happy.5 ~~ res_var*happy.5
happy.6 ~~ res_var*happy.6


'

```

Print the results

```{r}
model1 <- sem(linear_change_string, data = df_wide)
summary(model1)
```

