---
title: "cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Last Updated:
- Aug 10th, 2020 by Christopher

# Document Summary

1) Read in data
2) Select columns
3) Rename columns
4) Change to numeric
5) Select completed participants
6) Save data

## 1 Read in data

Use relative path

```{r}
df <- read.csv('../data/working_data.csv')
```


## 2 Select Columns

Select columns that I am interested in, which include Finished (Finished), computer id (Q157), session id (Q204), participant id (Q205), general self efficacy (SC0), and higher order need strength (SC1)

```{r}
df <- df %>%
  select(Finished,
         Q157,
         Q204,
         Q205,
         SC0,
         SC1)


```


## 3 Rename Columns

Change the column names so that they are easier to understand and use in scripts

```{r}
names(df) <- c('completed_survey',
               'computer_id',
               'session_id',
               'participant_id',
               'gse',
               'ho_need_strength')
```


## 4 Change to numeric

Most of the variables read in as factors or characters. I need to make them numeric in order to compute summary statistics and generate models.

```{r}

df[,] <- as.numeric(as.character(unlist(df[,])))

```


## 5 Select completed participants

Select participants who completed the entire study. Also remove practice survey attempts that I completed prior to the study (coded as participant id == 9999).

```{r}
df <- df %>%
  filter(completed_survey == 1) %>%
  filter(session_id != 'NA') %>%
  filter(participant_id < 999)

```


## 6 Save it

```{r}
write.csv(df, file = '../data/cleaned_data.csv', row.names = F)
```

